---
title: "R Notebook"
output: html_notebook
author: Xinyi Jin
date: October 2021
---

```{r}
library(Seurat)
library(dplyr)
library(SingleCellExperiment)
library(scater)
library(splatter)
library(reshape2)
library(ggplot2)
library(pROC)
library(ggforce) 
library(org.Hs.eg.db)
library(ggpubr)
library(scales)
library(limma)
library(foreach)
library(tictoc)
library(ggpubr)
library(clustree)

pbmc.data<- readRDS("pbmc_pooled_sex.Rds")
cm<- pbmc.data@assays$RNA@counts
inds<- as.vector(names(table(pbmc.data$individual)))

all.ind <- data.frame(individual = pbmc.data$individual)
all.ind$libsize <- 0
all.ind$pz = 0
all.ind$numgenes<- 0
all.ind$individual<- factor(all.ind$individual)
row.names(all.ind)<- colnames(pbmc.data)


for (i in inds){
  curr_cm <- pbmc.data[, pbmc.data$individual ==i ]@assays$RNA@counts
  all.ind[all.ind$individual == i, "libsize"] <- colSums(curr_cm)
  all.ind[all.ind$individual == i, "pz"] <-colMeans(curr_cm==0)
  all.ind[all.ind$individual == i, "numgenes"] <- colSums(curr_cm!=0)
}
cm<- pbmc.data@assays$RNA@counts
dim(cm)
num.cell.df<-data.frame(matrix(0, nrow=12, ncol=2))
colnames(num.cell.df)<- c("ids","num.cells")
num.cell.df$ids<- inds
for (i in 1:12){
  num.cell.df[i,2] <- ncol(pbmc.data[, pbmc.data$individual == inds[i]]@assays$RNA@counts)
}
num.cell.df<- num.cell.df[order(num.cell.df$num.cells),]
num.cell.df$ids<- factor(num.cell.df$ids, levels = num.cell.df[, "ids"])

 ggplot(data=num.cell.df, aes(x=ids, y=num.cells))+
  geom_bar(size=0.8,stat="identity", width=0.7)+
  xlab("individuals")+
  ylab("number of cells")+
  geom_text(aes(label=num.cells), vjust=-0.1, size=3)

ls<-ggplot(data= all.ind, aes(x = individual, y = libsize, fill = individual))+ 
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75) )+
  xlab("")+theme(legend.position = "none")+
  ylab("lbrary size")

pz<- ggplot(data= all.ind, aes(x=individual, y = pz, fill = individual))+ 
   geom_violin(draw_quantiles = c(0.25, 0.5, 0.75) )+
  theme(legend.position = "none")+ylab("proportion of zeros")+xlab("")

nc<-ggplot(data= all.ind, aes(x=individual, y = numgenes, fill =individual))+ 
   geom_violin(draw_quantiles = c(0.25, 0.5, 0.75) )+theme(legend.position = "none")+
  ylab("number of detected genes per cell")+xlab("")


ggarrange(nc, ls,pz, ncol = 1,  nrow = 3,legend="none", align="hv" )
ggsave(paste("Figure 4.1.1", "pdf",sep="."), width = 9, height=10)


ann.all<- readRDS("allpbmc.Rds")
numzero.genes <- rowSums(cm==0)
keep.genes <- numzero.genes < (ncol(cm)-10)
all.keep <- cm[keep.genes,]

ann.keep.all <- ann.all[keep.genes,]
cm.processed<- all.keep

# the above is the same as following 
#ann <- AnnotationDbi:::select(org.Hs.eg.db,keys=rownames(pbmc.data),
#                             columns=c("SYMBOL","ENTREZID","GENENAME"),
#                              keytype = "SYMBOL")
#m <- match(rownames(cm),ann$SYMBOL)
#ann <- ann[m,]
#numzero.genes <- rowSums(cm==0)
#keep.genes <- numzero.genes < (ncol(cm)-10)
#all.keep <- cm[keep.genes,]
#ann.keep.all <- ann[keep.genes,]
#cm.processed<- all.keep


pbmc <- CreateSeuratObject(counts = cm.processed,min.features = 500, 
                           min.cells = 10)
dim(pbmc) #15321 16687
```


```{r}
set.seed(9887)
pbmc <- SCTransform(pbmc, verbose = FALSE)
pbmc <- ScaleData(pbmc, verbose = FALSE)
pbmc <- RunPCA(pbmc, npcs = 50, verbose = FALSE)
ElbowPlot(pbmc,ndims=50)
```

```{r}

pbmc <- FindNeighbors(pbmc, dims = 1:30)
pbmc <- FindClusters(pbmc, resolution = c(0.8,0.9, 1, 1.1, 1.2, 1.3, 1.4))
table(Idents(pbmc))
table(pbmc$SCT_snn_res.1.4)
clustree(pbmc, prefix = "SCT_snn_res.")
```

```{r}
set.seed(98)
pbmc <- RunTSNE(pbmc, reduction = "pca", dims = 1:30)
Idents(pbmc)<- pbmc.data[,colnames(pbmc)]$predicted.celltype.l1
pdf("Figure4.2.1.pdf")
pbmc <- RunUMAP(pbmc, dims = 1:10)
plot(DimPlot(pbmc, reduction = "tsne",label=TRUE,label.size = 6, repel = TRUE)+
    scale_color_brewer(palette="Set1")
  )
dev.off()
```

# Background 
```{r}
library(tictoc)
library(foreach)
individuals<- unique(pbmc.data$individual)
# the list of true markers are downloded from 
# https://www.proteinatlas.org/humanproteome/celltype/blood+%26+immune+cells
files<-list("cell_type_category_rna_B-cells_Cell.tsv",
            "blood_cell_category_rna_naive_CD4T_elevated.tsv",
            "blood_cell_category_rna_naive_CD8T_elevated.tsv","",
            "cell_type_category_rna_Monocytes_Cell.tsv")

markers<- list()
unique.markers<- list()
all.ct<- c("B", "CD4 T", "CD8 T", "DC", "Mono", "NK", "other", "other T")
for (i in 1:8){
  if ( i %in% c(1, 2, 3, 5)){
    
  gene.df<- read.table(files[[i]], sep = '\t', header = TRUE)
  print(dim(gene.df))
  markers[[all.ct[i]]]<- gsub(x = c(unlist(gene.df$Gene),unlist(strsplit(gene.df$Gene.synonym, ","))), 
                      pattern=" ", replacement = "")
  unique.markers[[all.ct[i]]]<-gene.df$Gene
    
  }
}
for (ind in 1:length(individuals)){
  print(paste( individuals[ind], "=", as.character(unique(cmpbmc.data[, pbmc.data$individual == individuals[ind]]$sex))))
}

for (ind in 1:length(individuals)){
  cm<- pbmc.data[, pbmc.data$individual == individuals[ind]]@assays$RNA@counts
  dim(cm) 
  pbmc <- CreateSeuratObject(counts = cm,min.features = 500, min.cells = 5)
  cm<-pbmc@assays$RNA@counts
  numzero.genes <- rowSums(cm==0)
  keep.genes <- numzero.genes < (ncol(cm)-10)
  all.keep <- cm[keep.genes,]
  ann.keep.all <- ann.all[keep.genes,]
  cm.processed<- all.keep
  groups<-pbmc.data[row.names(cm.processed),colnames(cm.processed)]$predicted.celltype.l1
  table(groups)
  obs<-  computeObservation(cm.processed, factor(groups),statistic="os.t")
  
  os.t.res<- readRDS(paste("id_",individuals[ind],"_10kperm_os.t.Rds", sep=""))
pdf(paste("id",individuals[ind],"pvalues","pdf", sep="."))
  for (ct in 1:ncol(obs$obs.pval)){
    hist(os.t.res$obs.pval[,ct],
         main = paste("raw p-values by limma", "-", colnames(os.t.res$obs.pval.adj)[ct],"cells"), 
         xlab  = "raw p-values")
    
    hist(os.t.res$perm.pval[,ct],
         main = paste("raw p-values by permutation", "-", colnames(os.t.res$obs.pval.adj)[ct],"cells"), 
         xlab  = "raw p-values")
    
    hist(os.t.res$obs.pval.adj[,ct],
         main = paste("adjusted p-values by limma", "-", colnames(os.t.res$obs.pval.adj)[ct],"cells"), 
         xlab  = "adjusted p-values")
    
    hist(os.t.res$perm.pval.adj[,ct],
         main = paste("adjusted p-values by permutation", "-", colnames(os.t.res$obs.pval.adj)[ct],"cells"), 
         xlab  = "adjusted p-values")
    
    curr.ct.name=colnames(obs$obs.pval)[ct]
  curr_df<- data.frame(cbind("Amean"=obs$fit.cont.eb.obs$Amean, 
                      "coef"=obs$fit.cont.eb.obs$coefficients[,ct],
                      "trueMarkers"=rep(0, nrow(obs$obs.stat))
                      ))
      curr_df$Amean<-as.numeric(curr_df$Amean)
      curr_df$coef<-as.numeric(curr_df$coef)
      curr_df$trueMarkers<-factor(curr_df$trueMarkers, levels = c(0, 1))
      curr_df$trueMarkers[match(markers[[curr.ct.name]],row.names(obs$obs.pval))]<-1 
      limma.p<- ggplot(data=curr_df,#[curr_df$trueMarkers==1, ], 
                 aes(x =Amean, y=coef, color=trueMarkers, size=trueMarkers, shape=trueMarkers))+
        geom_point(alpha=0.4)+
        geom_hline(yintercept=0, size=0.5, color = "black" )+
        theme(legend.title=element_blank(),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="none",
         legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15))+
        xlab("Average log-expression")+ylab("log-fold-change")+
        ggtitle(paste("Marker genes for", curr.ct.name, "cells",sep=" "))+
        scale_size_manual(values=c(0.01,2))+
        scale_color_manual(values=c("black","red"))+
        scale_shape_manual(values=c(20, 17))+
        xlim(-1.5, 10)+ylim(-1.5, 5)
        ggsave(paste(paste("TrueDE_MAplot", "ind", individuals[ind],"stat_os.t",
                         "ct",curr.ct.name, sep="_"), "pdf", sep="."), width = 5, height=5)
      
      }
}
summary(os.t.res$result.table[[1]][match(markers[["B"]], row.names((os.t.res$obs.stat))),
                                   "logFC.g1"])
summary(os.t.res$result.table[[2]][match(markers[["CD4 T"]], row.names((os.t.res$obs.stat))), 
                                   "logFC.g2"])
summary(os.t.res$result.table[[3]][match(markers[["CD8 T"]], row.names((os.t.res$obs.stat))), 
                                   "logFC.g3"])
summary(os.t.res$result.table[[5]][match(markers[["Mono"]], row.names((os.t.res$obs.stat))), 
                                   "logFC.g5"])
```

## cell type poportions for each individual
```{r}
ct.df<-data.frame(matrix(0, nrow=12, ncol = 10))
colnames(ct.df)<- c("ids", "total.num.cells",
                    names(table(pbmc.data[,colnames(cm.processed)]$predicted.celltype.l1)))
for (i in 1:12){
  ct.df[i, "ids"]<-individuals[i]
  ct.df[i, "total.num.cells"]<- ncol(pbmc.data[, pbmc.data$individual == individuals[i]]@assays$RNA@counts)
  ct<- as.list(table(pbmc.data[,pbmc.data$individual == individuals[i]]$predicted.celltype.l1))
  for (j in 3:10){
    ct.df[i, colnames(ct.df)[j]]<- ifelse(colnames(ct.df)[j] %in% names(ct), 
                                          yes=ct[[colnames(ct.df)[j]]], no=0)
  }
}
ct.df<- ct.df[order(ct.df$total.num.cells),]
ct.df$ids<- factor(ct.df$ids, levels = ct.df$ids)



df<- melt(ct.df, id.vars =c("ids","total.num.cells"),
          variable.name = "cell.type", 
          value.name ="cell.type.number")


nums<-ggplot(data=df, aes(x=ids, fill = cell.type, y=cell.type.number, group=ids))+
  geom_bar(width=0.7, stat="identity")+
  xlab("individuals")+
  ylab("number of cells")+
  scale_fill_brewer(palette="Set1")+
  theme(
    axis.text.x  =element_text(size=6), 
        axis.title.x=element_text(size=10, vjust = 0.4),
        axis.title.y=element_text(size=10), 
        panel.spacing = unit(0.5, "lines"))+
  guides(fill=guide_legend(title="cell types",nrow = 4, ncol=2))

ct.prop<- cbind(ct.df[,1:2], prop.table(as.matrix(ct.df[,3:10]), 1)*100)
df<- melt(ct.prop, id.vars =c("ids","total.num.cells"),
          variable.name = "cell.type", 
          value.name ="cell.type.proportion")

prop<- ggplot(data=df, aes(x=cell.type, fill = ids, y=cell.type.proportion))+
  geom_bar(width=0.7, stat="identity", position=position_dodge())+
  xlab("cell types")+
  ylab("cell type proportion (%)")+
  scale_fill_brewer(palette="Set3")+
  theme(
        axis.title.x=element_text(size=10),
        axis.title.y=element_text(size=10), 
        panel.spacing = unit(0.5, "lines"))+
  guides(fill=guide_legend(title="individuals",nrow = 6, ncol=2))

ct.prop<- cbind(ct.df[,1:2],apply(round(prop.table(as.matrix(ct.df[,3:10]), 1)*100, 3), 
                                  FUN=paste, "%",sep="",MARGIN= c(1,2))
                )


ggarrange(nums, prop,heights = c(3,3), ncol = 1,  nrow = 2,
          legend="right", align="hv", labels = c("(A)", "(B)"), 
          vjust = 1.1, font.label = list(size=15))
ggsave(paste("Figure4.1.2", "pdf",sep="."), width = 8, height=7)
```


# Test statistics: ordinary t, moderated t and logFC*(1-p)
## this result is based on the log-transformed dataset (without normalisation)
```{r}


for (ind in 1:length(individuals)){

  cm<- pbmc.data[, pbmc.data$individual == individuals[ind]]@assays$RNA@counts
  dim(cm) # 32738  3402 (1789+1613) $32738  4907
  
  pbmc.ind <- CreateSeuratObject(counts = cm,min.features = 500, min.cells = 5)
  
  cm<-pbmc.ind@assays$RNA@counts
 
  numzero.genes <- rowSums(cm==0)
  keep.genes <- numzero.genes < (ncol(cm)-10)
  all.keep <- cm[keep.genes,]
  ann.keep.all <- ann.all[keep.genes,]
  cm.processed<- all.keep
  dim(cm.processed)
  n.cores=parallel::detectCores()-8
  perm.size = 10000
  groups<-pbmc.data[row.names(cm.processed),colnames(cm.processed)]$predicted.celltype.l1
  table(groups)
  
  all.res<- pPermMultipleStatistics(cm.processed, groups, perm.size = perm.size, n.cores=n.cores)
  os.t.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "os.t",
                           n.cores=n.cores, perm.arrays =all.res$os.t.stat )
  os.mot.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "os.modt",
                           n.cores=n.cores, perm.arrays =all.res$os.modt.stat )
  lfc.p.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "lfc.p",
                           n.cores=n.cores, perm.arrays =all.res$lfc.p.stat )
  
  # I saved all the objects locally
  #saveRDS(os.t.res, paste("id_",individuals[ind],"_10kperm_os.t.Rds", sep=""))
  #saveRDS(os.mot.res, paste("id_",individuals[ind],"_10kperm_os.modt.Rds", sep=""))
  #saveRDS(lfc.p.res, paste("id_",individuals[ind],"_10kperm_lfc.p.Rds", sep=""))
  #os.t.res<- readRDS(paste("id_",individuals[ind],"_10kperm_os.t.Rds", sep=""))
  #os.mot.res <- readRDS(paste("id_",individuals[ind],"_10kperm_os.modt.Rds", sep=""))
  #lfc.p.res<- readRDS(paste("id_",individuals[ind],"_10kperm_lfc.p.Rds", sep=""))

  stats.res<-list(os.t.res,os.mot.res,lfc.p.res ) 
  stats.name<- c("os.t","os.modt", "lfc.p")
  showed.name<- c("ordinary.t","moderated.t", "W statistic")
  if (exists("summary.stat") == FALSE ){
    cols.n<- c("id","cell.type","num.cells", "total.num.cells","test.name", 
                       "num.true.marker","num.upreg","TP", "precision", 
                       "sensitivity","F1", "overlap","TP.overlap","pvalue.cutoff",
                       "perm.size", "AUC","statistic")
      summary.res<- setNames(data.frame(matrix(ncol = length(cols.n), nrow = 0)),cols.n)
    }
  for (stat in 1:length(stats.res)){
    res<- stats.res[[stat]]
    obs<-  computeObservation(cm.processed, factor(groups),statistic=stats.name[stat])
    
   plt.title<-paste("summaryplots",individuals[ind],"stat",showed.name[stat], sep="_")
   # if summary plots required
   #auc<- summaryPlots(plt.title =plt.title , obs=obs, result= res,markers = markers,
   #                    statistic=showed.name[stat] )
   auc<- computeAUC(result= res,markers = markers)
   summary.stat<- computeComparison(res$obs.pval.adj, res$perm.pval.adj, 
                                       marker.genes=markers, ids = individuals[ind],
                                       unique.marker.genes = unique.markers, 
                             cell_num=as.vector(table(groups)), 
                             perm.size=10000, auc.values=auc,
                             statistic=stats.name[stat])
   summary.res<- rbind(summary.res, summary.stat)
   
  }
  
}
saveRDS(summary.res, "summary.res.12ind.Rds")


```


# Test statistics: ordinary t, moderated t and logFC*(1-p)
## this result is based on the cyclic loess normalised dataset
```{r}
library(tictoc)
library(foreach)
setwd("~/Desktop/Data Science Research Project/MDSresearchproject")
individuals<- unique(pbmc.data$individual)
ann.all<- readRDS("allpbmc.Rds")
files<-list("cell_type_category_rna_B-cells_Cell.tsv",
            "blood_cell_category_rna_naive_CD4T_elevated.tsv",
            "blood_cell_category_rna_naive_CD8T_elevated.tsv","",
            "cell_type_category_rna_Monocytes_Cell.tsv")

markers<- list()
unique.markers<- list()
all.ct<- c("B", "CD4 T", "CD8 T", "DC", "Mono", "NK", "other", "other T")
for (i in 1:8){
  if ( i %in% c(1, 2, 3, 5)){
    
  gene.df<- read.table(files[[i]], sep = '\t', header = TRUE)
  print(dim(gene.df))
  markers[[all.ct[i]]]<- gsub(x = c(unlist(gene.df$Gene),unlist(strsplit(gene.df$Gene.synonym, ","))), 
                      pattern=" ", replacement = "")
  unique.markers[[all.ct[i]]]<-gene.df$Gene
    
  }

}
for (ind in 1:length(individuals)){
  cm<- pbmc.data[, pbmc.data$individual == individuals[ind]]@assays$RNA@counts
  dim(cm) 
  
  pbmc.ind <- CreateSeuratObject(counts = cm,min.features = 500, min.cells = 5)
  
  cm<-pbmc.ind@assays$RNA@counts
 
  numzero.genes <- rowSums(cm==0)
  keep.genes <- numzero.genes < (ncol(cm)-10)
  all.keep <- cm[keep.genes,]
  ann.keep.all <- ann.all[keep.genes,]
  cm.processed<- all.keep
  dim(cm.processed)
  n.cores=parallel::detectCores()-8
  perm.size = 10000
  groups<-pbmc.data[row.names(cm.processed),colnames(cm.processed)]$predicted.celltype.l1
  table(groups)
  
  all.res<- pPermMultipleStatistics(cm.processed, groups, perm.size = perm.size, 
                                    n.cores=n.cores, normalise = TRUE)
  os.t.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "os.t",
                           n.cores=n.cores, perm.arrays =all.res$os.t.stat, 
                           normalise = TRUE )
  os.mot.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "os.modt",
                           n.cores=n.cores, perm.arrays =all.res$os.modt.stat,
                           normalise = TRUE )
  lfc.p.res<- runPermuation(cm.processed, group=groups, perm.size, statistic = "lfc.p",
                           n.cores=n.cores, perm.arrays =all.res$lfc.p.stat,
                           normalise = TRUE )
  
  #saveRDS(os.t.res, paste("id_",individuals[ind],"_10kperm_normalised_os.t.Rds", sep=""))
  #saveRDS(os.mot.res, paste("id_",individuals[ind],"_10kperm_normalised_os.modt.Rds", sep=""))
  #saveRDS(lfc.p.res, paste("id_",individuals[ind],"_10kperm_normalised_lfc.p.Rds", sep=""))

  #os.t.res<- readRDS(paste("id_",individuals[ind],"_10kperm_normalised_os.t.Rds", sep=""))
  #os.mot.res<- readRDS(paste("id_",individuals[ind],"_10kperm_normalised_os.modt.Rds", sep=""))
  #lfc.p.res<- readRDS(paste("id_",individuals[ind],"_10kperm_normalised_lfc.p.Rds", sep=""))

  
  stats.res<-list(os.t.res,os.mot.res,lfc.p.res ) 
  stats.name<- c("os.t","os.modt", "lfc.p")
  showed.name<- c("ordinary.t","moderated.t", "W statistic")

  if (exists("summary.stat") == FALSE ){
    cols.n<- c("id","cell.type","num.cells", "total.num.cells","test.name", 
                       "num.true.marker","num.upreg","TP", "precision", 
                       "sensitivity","F1", "overlap","TP.overlap","pvalue.cutoff",
                       "perm.size", "AUC","statistic")
      summary.res<- setNames(data.frame(matrix(ncol = length(cols.n), nrow = 0)),cols.n)
    }
  for (stat in 1:length(stats.res)){
    res<- stats.res[[stat]]
    obs<-  computeObservation(cm.processed, factor(groups),
                              statistic=stats.name[stat], normalise = TRUE)

   # if summary plots required
   #plt.title<-paste("normalised_summaryplots",individuals[ind],
   #                  "stat",showed.name[stat], sep="_")
   #auc<- summaryPlots(plt.title =plt.title , obs=obs, result= res,markers = markers,
   #                    statistic=showed.name[stat] )
   auc<- computeAUC(result= res,markers = markers)
   
   summary.stat<- computeComparison(res$obs.pval.adj, res$perm.pval.adj, 
                                       marker.genes=markers, ids = individuals[ind],
                                       unique.marker.genes = unique.markers, 
                             cell_num=as.vector(table(groups)), 
                             perm.size=10000, auc.values=auc,
                             statistic=stats.name[stat])
    summary.res<- rbind(summary.res, summary.stat)

  
  }
  
}
saveRDS(summary.res,"summary.res.normalised.12ind.Rds")
```

# plots 
```{r}


summary.res$category<- paste(summary.res$test.name,summary.res$statistic)
summary.res$cell.proportion<- 100*summary.res$num.cells/summary.res$total.num.cells

#summary.res<- readRDS("summary.res.12ind.Rds")
df2<- summary.res[summary.res$cell.type %in% c("B", "CD4 T", "CD8 T", "Mono"),]

df2[df2$category=="limma lfc.p" , "category"]<-"limma os.t"

df2$cell.type<- factor(df2$cell.type, levels = c("Mono","B", "CD8 T","CD4 T"))
df2$category<- factor(df2$category)
df2$category<- gsub(df2$category, pattern = "os.t", replacement = "ordinary.t")
df2$category<- gsub(df2$category, pattern = "os.modt", replacement = "moderated.t")
df2$category<- gsub(df2$category, pattern = "lfc.p", replacement = "W")
df2$category<- factor(df2$category,
                      levels= c("limma ordinary.t","permutation ordinary.t",
                                "limma moderated.t","permutation moderated.t",
                                "permutation W"))
df2$id<- factor(df2$id, levels=ct.df[, "ids"])

clrs<-c('lightsteelblue2','navajowhite2', "steelblue2","darkgoldenrod1","green4")
# colors corresponds to "limma ordinary.t","permutation ordinary.t", limma moderated.t","permutation moderated.t","permutation W" 
```

Overall sensitivity, precision and F1 score for different cell types. Each dot represents an individual, and the result is based on log-transformed data
```{r}

# figure 4.3.2
sensitivity<-ggplot(data = df2,
                    aes(x = cell.type, y = sensitivity, color=category))+
  geom_boxplot(size=0.2, width=0.8,outlier.shape = NA)+
  geom_point(position=position_jitterdodge(jitter.width=0, jitter.height=0), 
             size=1,  alpha=0.9)+
  theme(legend.title=element_blank(),
        axis.title.x=element_text(size=10),
        axis.title.y=element_text(size=10), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
         legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15),
        strip.text = element_text(size = rel(1.3)))+
  guides(color = guide_legend(nrow=1, ncol =length(clrs)))+
  xlab("cell types")+
  ylab("sensitivity")+ 
  scale_y_continuous(limits=c(0,0.45), oob = rescale_none)+
  scale_color_manual(values=clrs)

precision<-ggplot(data = df2,
                    aes(x = cell.type, y = precision, color=category))+
  geom_boxplot(size=0.2, width=0.8,outlier.shape = NA)+ #position = "fill", )+
  geom_point(position=position_jitterdodge(jitter.width=0, jitter.height=0), 
             size=1,  alpha=0.9)+
  theme(legend.title=element_blank(),
        axis.title.x=element_text(size=10),
        axis.title.y=element_text(size=10), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
         legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15),
        strip.text = element_text(size = rel(1.3)))+
  guides(fill = guide_legend(nrow=1, ncol =length(clrs)))+
  xlab("cell types")+
  ylab("precision")+ 

  scale_y_continuous(limits=c(0,0.5), oob = rescale_none)+
  scale_color_manual(values=clrs)

ggarrange(sensitivity+xlab(""), precision, heights = c(3,3), ncol = 1,  nrow = 2,
          legend="top", common.legend = T,align="hv", labels = c("(A)","(B)"))
ggsave(paste("Figure4.3.2", "pdf",sep="."), width = 13, height=8)

```

Performance on finding marker genes for B, Monocytes, CD4 T and CD8 T cells for each individual
```{r}

# figure 4.3.3
sensitivity<-ggplot(data = df2,
                    aes(x = cell.type, y = sensitivity,fill =category))+
  geom_bar(width=0.7, stat="identity", position=position_dodge(), color="black")+
  facet_wrap(~id, nrow=3, ncol=4, scales = 'free_x')+
  theme(legend.title=element_blank(),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
        legend.text=element_text(margin=margin(r=0.2,unit="cm"), size=13),
        strip.text = element_text(size = rel(1.3)))+
  guides(fill = guide_legend(nrow=1, ncol =length(clrs)))+
  xlab("cell types")+
  ylab("sensitivity")+ 
  scale_y_continuous(limits=c(0,0.45), oob = rescale_none)+
  scale_fill_manual(values=clrs)

precision<-ggplot(data = df2,
                  aes(x = cell.type, y = precision,fill =category))+
  geom_bar(width=0.7, stat="identity",position=position_dodge(), color="black")+
  facet_wrap(~id, nrow=3, ncol=4, scales = 'free_x')+
  theme(legend.title=element_blank(),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
        legend.text=element_text(margin=margin(r=0.2,unit="cm"), size=13),
        strip.text = element_text(size = rel(1.3)))+
  guides(fill = guide_legend(nrow=1, ncol =length(clrs)))+
  xlab("cell types")+
  ylab("precision")+ 
  scale_y_continuous(limits=c(0,0.5), oob = rescale_none)+
  #scale_fill_brewer(palette="Set1")
  scale_fill_manual(values=clrs)

ggarrange(sensitivity+xlab(""), precision, heights = c(3,3), ncol = 1,  nrow = 2,
          legend="bottom", common.legend = T,align="hv")
ggsave(paste("Figure4.3.3", "pdf",sep="."), width = 12, height=8)
```

### Summary plot for ordinary t with original data, ordinary.t with normalised data and treat-t (0.1)
```{r}


summary.normalised<- readRDS("summary.res.normalised.12ind.Rds")
df<- summary.normalised[summary.normalised$cell.type %in% c("B", "CD4 T", "CD8 T", "Mono"),]
df$cell.type<- factor(df$cell.type, levels = c("Mono","B", "CD8 T","CD4 T"))
df$threshold <- 0
df$category<-paste(df$test.name, "normalised", df$statistic)
df$cell.proportion<- 100*df$num.cells/df$total.num.cells
df<- df[df$category %in% c("limma normalised os.t", "permutation normalised os.t"), ]


summary.org<- readRDS("summary.res.12ind.Rds")
df1<- summary.org[summary.org$cell.type %in% c("B", "CD4 T", "CD8 T", "Mono"),]
df1$cell.type<- factor(df1$cell.type, levels = c("Mono","B", "CD8 T","CD4 T"))
df1$threshold <- 0
df1$category<-paste(df1$test.name,df1$statistic)
df1$cell.proportion<- 100*df1$num.cells/df1$total.num.cells
df1<- df1[df1$category %in% c("limma os.t", "permutation os.t"), ]
df1$category<- gsub(df1$category, pattern = "os.t", replacement = "ordinary.t")


df.tt<- rbind(df1, df)
df.tt[df.tt$TP==0, c("precision","F1","sensitivity")]=0
df.tt$category<- gsub(x=df.tt$category, pattern = "os.t",replacement = "ordinary.t")
df.tt$category<-factor(df.tt$category, 
                       levels=c("limma ordinary.t","permutation ordinary.t",
                                "limma normalised ordinary.t",
                                "permutation normalised ordinary.t"))

library(RColorBrewer)
color_lst = c(brewer.pal(name="Paired", n = 4))
sens<-ggplot(data = df.tt,
                    aes(x = cell.type, y = sensitivity, fill=category))+
  geom_boxplot(outlier.shape = NA, position=position_dodge())+
  geom_point(position=position_jitterdodge(),size=0.8)+
  theme(legend.title=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_text(size=15), 
    panel.spacing = unit(0.5, "lines"), 
    legend.position="top",
    legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15),
    strip.text = element_text(size = rel(1.3)))+
  guides(fill = guide_legend(nrow=2, ncol=2))+
  xlab("cell types")+
  ylab("sensitivity")+ 
  scale_fill_manual(values=color_lst)+
  scale_y_continuous(limits=c(0,0.6), oob = rescale_none)

prec<-ggplot(data = df.tt,
                  aes(x = cell.type, y = precision, fill=category))+
  geom_boxplot(outlier.shape = NA, position=position_dodge())+
  geom_point(position=position_jitterdodge(),size=0.8)+
  theme(legend.title=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_text(size=15), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
        legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15),
        strip.text = element_text(size = rel(1.3)))+
  guides(fill = guide_legend(nrow=2, ncol=2))+
  scale_y_continuous(limits=c(0,0.5), oob = rescale_none)+
  xlab("cell types")+
  scale_fill_manual(values=color_lst)+
  ylab("precision")
F1<-ggplot(data = df.tt,
                    aes(x = cell.type, y = F1, fill=category))+
  geom_boxplot(#size=0.5, width=0.9,
    outlier.shape = NA, position=position_dodge())+
  geom_point(position=position_jitterdodge(),size=0.8)+
    theme(legend.title=element_blank(),
          axis.title.x=element_text(size=15),
          axis.title.y=element_text(size=15), 
          panel.spacing = unit(0.5, "lines"), 
          axis.text.x = element_text(size=15),
          legend.position="top",
          legend.text=element_text(margin=margin(r=0.3,unit="cm"), size=15),
          strip.text = element_text(size = rel(1.3)))+
    guides(fill = guide_legend(nrow=2, ncol=2))+
    xlab("cell types")+
    ylab("F1 score")+ 
   scale_fill_manual(values=color_lst)+
  scale_y_continuous(limits=c(0,0.35), oob = rescale_none)

ggarrange(sens+xlab(""), prec+xlab(""), F1, heights = c(3,3,3), 
          ncol = 1,  nrow = 3,
          legend="top", common.legend = T,align="hv"
          )

ggsave(paste("Summary_of_three_stat_PBMC", "pdf",sep="."), width = 8, height=6)


```

### cell types without true marker genes 
```{r}

summary.res$category<- paste(summary.res$test.name,summary.res$statistic)
summary.res$cell.proportion<- 100*summary.res$num.cells/summary.res$total.num.cells

#summary.res<- readRDS("summary.res.12ind.Rds")
df3<- summary.res
df3[df3$category=="limma lfc.p" , "category"]<-"limma os.t"

df3$cell.type<- factor(df3$cell.type, 
                       levels = c("DC","other","other T",
                                  "Mono","B","NK","CD8 T","CD4 T")
                       )

df3$category<- factor(df3$category)
df3$category<- gsub(df3$category, pattern = "os.t", replacement = "ordinary.t")
df3$category<- gsub(df3$category, pattern = "os.modt", replacement = "moderated.t")
df3$category<- factor(df3$category,
                      levels= c("limma ordinary.t","permutation ordinary.t",
                                "limma moderated.t","permutation moderated.t",
                                "permutation lfc.p"))



m<-ggplot(data = df3,
                    aes(x = cell.type, y = num.upreg,
                        fill =category))+
  geom_bar(width=0.7, stat="identity", position=position_dodge())+
  facet_wrap(~id, nrow=3, ncol=4)+
  theme(legend.title=element_blank(),
        axis.title.x=element_text(size=15),
        axis.title.y=element_text(size=15), 
        panel.spacing = unit(0.5, "lines"), 
        legend.position="top",
        legend.text=element_text(margin=margin(r=0.2,unit="cm"), size=13),
        strip.text = element_text(size = rel(1.3)))+
  xlab("cell types")+
  ylab("number of detected marker genes")+ 
  scale_fill_manual(values=clrs)
ggsave(paste("Figure 4.3.6", "pdf",sep="."), width = 14, height=5)
```

### heatmap of top-20 most significant genes
```{r}
setwd("/Volumes/Expansion/Data Science Research Project/MDSresearchproject/RealDataAnalysis/normalised_v")
os.t.normalised<- readRDS(paste("id_","689_690","_10kperm_normalised_os.t.Rds", sep=""))

cm<- pbmc.data[, pbmc.data$individual == "689_690"]@assays$RNA@counts
dim(cm) 
pbmc.ind <- CreateSeuratObject(counts = cm,min.features = 500, min.cells = 5)
cm<-pbmc.ind@assays$RNA@counts
numzero.genes <- rowSums(cm==0)
keep.genes <- numzero.genes < (ncol(cm)-10)
all.keep <- cm[keep.genes,]
ann.keep.all <- ann.all[keep.genes,]
cm.processed<- all.keep
dim(cm.processed)
n.cores=parallel::detectCores()-8
perm.size = 10000
groups<-pbmc.data[row.names(cm.processed),colnames(cm.processed)]$predicted.celltype.l1
os.t.normalised<- runPermuation(cm.processed, group=groups, perm.size, statistic = "os.t",
                           n.cores=n.cores, perm.arrays =all.res$os.t.stat, 
                           normalise = TRUE )
# I saved all the objects, so just read it
# os.t.normalised<- readRDS(paste("id_",individuals[7],"_10kperm_normalised_os.t.Rds", sep=""))

pbmc.ind <- SCTransform(pbmc.ind, verbose = FALSE)
pbmc.ind <- RunPCA(pbmc.ind, npcs = 50, verbose = FALSE)


for (i in 1:8){
  tb<- os.t.normalised$result.table[[i]]
  sig_genes<- tb[tb$limma.upreg==1 & tb$permutation.upreg==1, ]
  sig_genes<-sig_genes[order(sig_genes$limma.adj.pvalue, -sig_genes[,1],sig_genes$perm.adj.pvalue), ]
  assign(paste("sig",colnames(os.t.normalised$perm.pval)[i], sep="."), row.names(sig_genes)[1:20])
} 


table(`sig.CD4 T` %in% markers[["CD4 T"]])
table(`sig.CD8 T` %in% markers[["CD8 T"]])
table(sig.Mono %in% markers[["Mono"]])
table(sig.B %in% markers[["B"]])
genes<-c(`sig.CD4 T`, `sig.CD8 T`, sig.NK, sig.Mono, sig.B)
keep.cells<-colnames(pbmc.ind)[pbmc.data[,colnames(pbmc.ind)]$predicted.celltype.l1 %in% c("Mono","B","CD8 T","CD4 T","NK")]
pbmc.ind<-pbmc.ind[, keep.cells]
Idents(pbmc.ind)<- pbmc.data[,colnames(pbmc.ind)]$predicted.celltype.l1
table(Idents(pbmc.ind))
DoHeatmap(pbmc.ind,features = genes,group.by="ident",label = T)+
  guides(color = FALSE)
```

```{r}
print(sessionInfo())
```

